### 2015-05-31 abstract.tex
Put things into git

\documentclass[11pt]{article}

\usepackage{amsmath,amssymb,fancyhdr}
\usepackage[margin=1in,includehead]{geometry}

\begin{document}

  \begin{center}
    \Large\bf\boldmath Determining the Structure of Length-$k$ Steenrod Operations as $A(r)$-Modules
  \end{center}

  The Steenrod Algebra $\mathcal{A}$ is the algebra of stable natural endomorphisms of the $\mathbb{Z}/2$-cohomology functor; it is generated by elements $Sq^{2^i}$.  Let $\mathcal{A}(k)$ be the subalgebra generated by the $Sq^{2^i}$ for $i\leq k$.  Consider the modules $L(k)$ spanned by sequences of Steenrod operations of length $k$.  Welcher proved that $L(k)$ is a free module over $\mathcal{A}(k-1)$.  We are interested in finding the structure of $L(k)$ as an $\mathcal{A}(r)$-module for any $r$.  We conjecture that $L(k)$ is built as an $\mathcal{A}(r)$-module out of $\mathcal{A}(r)/\!/\!\mathcal{A}(r-k)$, in the sense that it has an increasing filtration with quotients isomorphic to $\mathcal{A}(r)/\!/\!\mathcal{A}(r-k)$, and present partial results towards that claim.  In addition, we prove some interesting commutation relations in the Steenrod algebra relating to representations of Steenrod Algebra elements in Wood's $Z$-basis.

\end{document}