### 2015-05-31 snippets/free-module.tex

\documentclass{article}

\usepackage{amsmath,amssymb,amsfonts,amsthm}

\newcommand{\A}{\mathcal{A}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\Sq}{\mathrm{Sq}}

\newtheorem{prop}{Proposition}
\newtheorem{lem}{Lemma}

\begin{document}
  For convenience let $\bar{L}(k)=\bigcup_{i\leq k}L(i)$.

  \begin{prop}
    $\bar{L}(k)//A(n-k)$ is a free module over $A(n)//A(n-k)$, for any $n\geq k$.
  \end{prop}

  In particular, we will show that $\bar{L}(k)//A(n-k)$ has $A(n)//A(n-k)$-basis
  \[
    B=\{\Sq^I\mid I=(2^{n+1}j_1, 2^nj_2,\ldots, 2^{n-k+2}j_k), j_1\geq j_2\geq\cdots\geq j_k\geq 0\}.
  \]

  First, let us show that the module generated by $B$ is free.  Here we follow a similar argument of Welcher.  Let $t$ be the top class of $A(n)//A(n-k)$.  Then I claim that for any $A(n)//A(n-k)$-module $M$, the submodule generated by $m\in M$ is free if $tm\neq 0$.  % This depends on something in Margolis that I need to find.

  % Now I claim that the top class is
  % \[t=\Sq^{1+n 2^{n+1}}\Sq^{1+(n-1)2^n}\cdots\Sq^{1+(n-k+1)2^{n-k+2}} + t_0\Sq^1 + t_1\Sq^2 + \cdots + t_{n-k}\Sq^{2^{n-k}},\]
  % for some elements $t_i$.
  % By looking at Wood's Z-basis and the Adem relations it suffices to prove that
  % \[\Sq^{2^{n+1}-2^n}\Sq^{2^{n+1}-2^{n-1}}\cdots\Sq^{2^{n+1}-1}=\Sq^{1+n 2^{n+1}} + t_0\Sq^1 + t_1\Sq^2 + \cdots + t_{n-1}\Sq^{2^{n-1}},\]
  % again for some elements $t_i$.  To see this, note first that the two elements are in the right dimension.  Now since Wood's basis is a basis, the left side is nonzero in $A(n)//A(n-k)$, so it may be written as an element of $L(1)$ plus an element of $A(\Sq^i)_{i\leq n-1}$; the only possible nontrivial choice of the former is $\Sq^{1+n 2^{n+1}}$, and the latter by definition may be written in the form of the remaining terms.

  Then it suffices to show that $tB$ is linearly independent.  To do this, we consider only the top terms in the left lexicographic ordering, and show that no two are the same.  Following Welcher, we let $\equiv$ mean congruent modulo both lower terms and $A(\Sq^i)_{i\leq n-k}$.  In particular, we will show that if $K=(1+n 2^{n+1},1+(n-1)2^n,\ldots,1+(n-k+1)2^{n-k+2})$, and $\Sq^I\in B$, then $t\Sq^I\equiv\Sq^{I+K}$.  Clearly none of these are the same, so $tB$ will then be linearly independent.

  To show this, recall that in the Wood Z-basis, the top class $t$ is represented by the element
  \[\Sq^{2^{n+1}-2^n}\cdots\Sq^{2^{n+1}-1}\Sq^{2^n-2^{n-1}}\cdots\Sq^{2^{n-k+2}-2^{n-k+1}}\cdots\Sq^{2^{n-k+2}-1}.\]
  Then we get

  % \begin{lem} %not true!
  %   Let $t_{n,k}=\Sq^{1+n 2^{n+1}}\Sq^{1+(n-1)2^n}\cdots\Sq^{1+(n-k+1)2^{n-k+2}}$, with $t_{n,k}=1$ if $k\leq0$.  Then $t_{n,k}\Sq^{c 2^{n+1}} = \Sq^{c 2^{n+1} + 1 + n2^{n+1}}t_{n-1,k-1}$.
  % \end{lem}
  % \begin{proof}
  %   Fix $n-k$ and induct on $n$.  If $k=1$, we wish to show that
  %   \[\Sq^{1+n2^{n+1}}\Sq^{c2^{n+1}}\equiv\Sq^{1+n2^{n+1}+c2^{n+1}}

  % Next, let us show that $B$ has the appropriate size.  We use the Poincar\'e series $v(t)=\sum t^i \dim V_i$ of a graded vector space $V$; multiplication of tensor products corresponds to tensor products of graded vector spaces.  Let $l(t)$, $a(t)$, and $b(t)$ be the Poincar\'e series of $L(k)//A(n-k)$, $A(n)//A(n-k)$, and the space spanned by $B$, respectively.

  % By consideration of Wood's Z-basis, we get that
  % \[
  %   a(t) = \prod_{\substack{k+2\leq i\leq n+1\\ 0\leq j<i}} \left(1+t^{2^i-2^j}\right).
  % \]

\end{document}