### 2015-05-31 snippets/set-of-representatives.tex

\documentclass{article}

\usepackage{amsmath,amssymb,amsfonts,amsthm}

\newcommand{\A}{\mathcal{A}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\Sq}{\mathrm{Sq}}

\newtheorem{prop}{Proposition}
\newtheorem{lem}{Lemma}

\begin{document}

  For convenience let $\bar{L}(k)=\bigcup_{i\leq k}L(i)$.

  \begin{prop}
    % There is a set of representatives for a basis of $\A(n)//\A(n-k)$ (in the sense of the above note), all of which lie in $\bar{L}(k)$.
    Any element $t$ of $\A(n)$ may be written as $x + yz$ where $x\in \bar{L}(k)$, $y\in \A$, and $z\in\A(n-k)\setminus \F_2$ for any $k\geq 0$.  Equivalently, $\A(n)\subset \bar{L}(k) + \A(\Sq^i)_{i\leq k}$, where we take $\A(\Sq^i)_{i\leq k}$ to be the left ideal of $\A$ generated by $\{\Sq^i\}_{i\leq k}$, for any $k\geq 0$.
  \end{prop}

  \begin{proof}
    It suffices to prove the proposition on some basis for $\A(n)$.  Recall that in Wood's Z-basis of the Steenrod algebra, a basis for $\A(n)$ consists of all strings of squares selected from~\eqref{eq:woodz} and multiplied in the given order:
    \begin{equation} \label{eq:woodz}
      \Sq^{2^{n+1}-2^n}\Sq^{2^{n+1}-2^{n-1}}\cdots\Sq^{2^{n+1}-1} \Sq^{2^n-2^{n-1}}\cdots\Sq^{2^n-1}\cdots\Sq^2\Sq^3\Sq^1
    \end{equation}

    We induct on the number of squares selected.  The empty product is $1$, which is in $\bar{L}(k)$, so we are done.  So now suppose the proposition holds on all elements of the Wood Z-basis of length at most $l-1$, and let $t$ be an element of length $l$.  Then, for some $m=2^j-2^i$, $i<j\leq n+1$, we may write $t=\Sq^m\left(x+yz\right)$ where $x$, $y$, and $z$ are as in the statement of the proposition.

    Now $\Sq^myz$ is clearly of the prescribed form, since $\Sq^my$ is in $\A$.  If $x\in\bar{L}(k-1)$, then $\Sq^m x\in \bar{L}(k)$ (since application of the Adem relations cannot increase the length of a monomial), so we may suppose $x\in L(k)$.

    Write $x$ in admissible form as $\Sq^{i_1}\Sq^{i_2}\cdots\Sq^{i_k}$.  If $\Sq^m\Sq^{i_1}\Sq^{i_2}\cdots\Sq^{i_k}$ is already in admissible form, then we have that $2^ki_k\leq 2^{k-1}i_{k-1}\leq\cdots\leq2i_1\leq m<2^{n+1}$, so in particular, $i_k<2^{n-k+1}$, so $i_k\in\A(\Sq^i)_{i\leq k}$, so in fact the entire product $\Sq^mx$ is in $\A(\Sq^i)_{i\leq k}$.  Then we may apply the Adem relations at least once, to get
    \[\Sq^m\Sq^{i_1}\Sq^{i_2}\cdots\Sq^{i_k} = a_0\Sq^{m+i_1}\Sq^{i_2}\cdots\Sq^{i_k} + \sum_{1\leq j\leq \frac{m}2}a_j\Sq^{m+i_1-j}\Sq^j\Sq^{i_2}\cdots\Sq^{i_k};\]
    the first term we know is in $L(k)$.  For each of the remaining terms, if we may repeat the same process, ignoring the first factor; either the product is in $\A(\Sq^i)_{i\leq k}$, or it may be further reduced by the Adem relations.  Finally, if we apply the Adem relations $k$ times, the rightmost factor must be $\Sq^i$ for $i\leq \frac{m}{2^k}<2^{n-k+1}$, so similarly the product lies in the ideal $\A(\Sq^i)_{i\leq k}$.  Then $\Sq^mx$ may be written in the desired form, so we are done.
  \end{proof}

\end{document}